Effect of quality-adjusted production function inputs

### 2015-05-14 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Lots of progress using formulas.

<<setup, echo=FALSE, message=FALSE, eval=TRUE>>=
require(EconData)
require(EconModels)
require(dplyr)
require(reshape2)
require(ggplot2)

@@

<<Data>>=
countries <- c("PT", "UK")
dataPT <- filter(IST, Country == "PT")
dataUK <- filter(Leeds, Country == "UK")

# Create some pre-econometric ratios
# Unadjusted variables
dataPT <- transform(dataPT, kPerGDP.unadj = dataPT[[vars$PT$capital$unadjusted]]/dataPT[[vars$PT$response$unadjusted]])
dataPT <- transform(dataPT, kPerGDP.unadj = iKstkS.L/iGDP)
dataPT <- transform(dataPT, kPerGDP.unadj = as.name(vars$PT$capital$unadjusted)/iGDP)
dataPT <- transform(dataPT, kPerGDP.unadj = eval(substitute(as.name(vars$PT$capital$unadjusted)))/iGDP)
dataPT <- transform(dataPT, kPerGDP.unadj = eval(quote(as.name(vars$PT$capital$unadjusted)))/iGDP)

dataPT <- within(filter(IST, Country == "PT"), {
  kPerGDP.unadj <- iKstkS.L / iGDP # Works
#   kPerGDP.unadj <- as.name(extract_rhs(forms$PT$unadjusted, 1)) / as.name(extract_lhs(forms$PT$unadjusted, 1)) # Fails
})

dataPT <- transform(dataPT, kPerGDP.unadj = eval(k_PT_unadj)/iGDP)
dataPT <- transform(dataPT, lPerGDP.unadj = eval(l_PT_unadj)/iGDP)
dataPT <- transform(dataPT, ePerGDP.unadj = eval(e_PT_unadj)/iGDP)
dataUK <- transform(dataUK, kPerGDP.unadj = eval(k_UK_unadj)/iGDP)
dataUK <- transform(dataUK, lPerGDP.unadj = eval(l_UK_unadj)/iGDP)
dataUK <- transform(dataUK, ePerGDP.unadj = eval(e_UK_unadj)/iGDP)
# Quality-adjusted variables
dataPT <- transform(dataPT, kPerGDP.adj = eval(k_PT_adj)/iGDP)
dataPT <- transform(dataPT, lPerGDP.adj = eval(l_PT_adj)/iGDP)
dataPT <- transform(dataPT, ePerGDP.adj = eval(e_PT_adj)/iGDP)
dataUK <- transform(dataUK, kPerGDP.adj = eval(k_UK_adj)/iGDP)
dataUK <- transform(dataUK, lPerGDP.adj = eval(l_UK_adj)/iGDP)
dataUK <- transform(dataUK, ePerGDP.adj = eval(e_UK_adj)/iGDP)

# Now create the historicalData data frame
historicalData <- plyr::rbind.fill(dataPT, dataUK)
@

<<Formulas>>=
# Set up the formulas for each type of analysis.
forms <- 
  list( 
    PT = c(unadjusted = build_formula("iGDP", list("iKstkS.L", "iL", "iXpMP", "iYear")),
             adjusted = build_formula("iGDP", list("iKservS.L", "ihLest", "iUMP", "iYear"))),
    UK = c(unadjusted = build_formula("iGDP", list("iKservO.WwithRD", "iL", "iXp", "iYear")),
             adjusted = build_formula("iGDP", list("iKservO.WwithRD", "ihLest", "iU", "iYear")))
  )

@

@@

\begin{table} \caption{Unadjusted time-series variables.} 
\label{tab:unadjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal           & UK                 \\
      \midrule
      Output      & \Sexpr{deparse(lhs(forms$PT$unadjusted))}   & \Sexpr{deparse(lhs(forms$UK$unadjusted))}   \\
      Capital     & \Sexpr{extract_rhs(forms$PT$unadjusted, 1)} & \Sexpr{extract_rhs(forms$UK$unadjusted, 1)} \\
      Labor       & \Sexpr{extract_rhs(forms$PT$unadjusted, 2)} & \Sexpr{extract_rhs(forms$UK$unadjusted, 2)} \\
      Energy      & \Sexpr{extract_rhs(forms$PT$unadjusted, 3)} & \Sexpr{extract_rhs(forms$UK$unadjusted, 3)} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{Quality-adjusted time series variables.} 
\label{tab:quality_adjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal         & UK               \\
      \midrule
      Output      & \Sexpr{deparse(lhs(forms$PT$adjusted))}   & \Sexpr{deparse(lhs(forms$UK$adjusted))}   \\
      Capital     & \Sexpr{extract_rhs(forms$PT$adjusted, 1)} & \Sexpr{extract_rhs(forms$UK$adjusted, 1)} \\
      Labor       & \Sexpr{extract_rhs(forms$PT$adjusted, 2)} & \Sexpr{extract_rhs(forms$UK$adjusted, 2)} \\
      Energy      & \Sexpr{extract_rhs(forms$PT$adjusted, 3)} & \Sexpr{extract_rhs(forms$UK$adjusted, 3)} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<Historical Data, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Historical data.", warning=FALSE>>=
unadjVars_PT <- all.vars(forms$PT$unadjusted)[1:4]
unadjVars_UK <- all.vars(forms$UK$unadjusted)[1:4]
commonVars <- c("Year", "iYear", "Country")
unadjData_PT <- historicalData[ , unique(c(unadjVars_PT, commonVars))] %>% filter(Country=="PT")
unadjData_UK <- historicalData[ , unique(c(unadjVars_UK, commonVars))] %>% filter(Country=="UK")
unadjData <- plyr::rbind.fill(unadjData_PT, unadjData_UK)
unadjData$type <- "Unadjusted"
adjVars_PT <- all.vars(forms$PT$adjusted)[1:4]
adjVars_UK <- all.vars(forms$UK$adjusted)[1:4]
adjData_PT <- historicalData[ , unique(c(adjVars_PT, commonVars))] %>% filter(Country=="PT")
adjData_UK <- historicalData[ , unique(c(adjVars_UK, commonVars))] %>% filter(Country=="UK")
adjData <- plyr::rbind.fill(adjData_PT, adjData_UK)
adjData$type <- "Quality-adjusted"
trimmedData <- plyr::rbind.fill(unadjData, adjData)

mdata <- melt(trimmedData, 
              measure.vars = unique(c(unadjVars_PT, unadjVars_UK, adjVars_PT, adjVars_UK)), 
              variable.name = "variable",
              value.name = "value")

# Relevel factors before plotting
mdata$Country <- relevelFactor(as.factor(mdata$Country), c("PT", "UK"))
mdata$type <- relevelFactor(as.factor(mdata$type), c("Unadjusted", "Quality-adjusted"))
mdata$variable <- relevelFactor(as.factor(mdata$variable), unique(c(unadjVars_PT, adjVars_PT, unadjVars_UK, adjVars_UK)))

historicalPlot(mdata, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = Country ~ type, 
               line_types = c(1,2,4,3,2,4,3,2,3,2,3)) + 
  scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
  coord_cartesian(ylim=c(0.5, 13)) +
  xlab("") + 
  ylab("Indexed (1960=1)")
@

<<Pre-econometric graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Pre-econometric data.", warning=FALSE>>=
# Melt the data
mdata_preeconometric <- melt(historicalData, 
                             measure.vars = c("kPerGDP.unadj", "lPerGDP.unadj", "ePerGDP.unadj",
                                              "kPerGDP.adj", "lPerGDP.adj", "ePerGDP.adj"),
                             variable.name = "variable",
                             value.name = "value")
# Add a column to say whether the variable is Unadjusted or Quality-adjusted based on variable name
mdata_preeconometric$adjusted <- lapply(mdata_preeconometric$variable, function(x) {
  suffix <- unlist(strsplit(as.character(x), split="[.]")) [2]
  if (suffix == "unadj"){
    return("Unadjusted")
  } else if (suffix == "adj"){
    return("Quality-adjusted")
  } else {
    stop(paste("Unknown suffix:", suffix))
  }
})
# Fix levels
mdata_preeconometric$adjusted <- relevelFactor(as.factor(as.character(mdata_preeconometric$adjusted)), 
                                               c("Unadjusted", "Quality-adjusted"))
mdata_preeconometric$Country <- relevelFactor(as.factor(mdata_preeconometric$Country), 
                                              c("PT", "UK"))
# Now make the plot
historicalPlot(mdata_preeconometric, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = Country ~ adjusted, 
               line_types = c(2,4,3, 2,4,3)) + 
  xlab("") + 
  ylab("Indexed (1960=1)")
@

@@

  models <- list()
  for ()

>>>>

<<setup, echo=FALSE, message=FALSE, eval=TRUE>>=
require(EconData)
require(EconModels)
require(dplyr)
require(reshape2)
require(ggplot2)
require(mosaic)

@@

<<Formulas>>=
# Set up the formulas for each type of analysis.
# These formulas contain energy. 
# We can use keep_summands to drop the energy term (3rd summand on the rhs)
# later if needed.
formulas <- 
  list( 
    PT = c(unadjusted = build_formula("iGDP", list("iKstkS.L", "iL", "iXpMP", "iYear")),
             adjusted = build_formula("iGDP", list("iKservS.L", "ihLest", "iUMP", "iYear"))),
    UK = c(unadjusted = build_formula("iGDP", list("iKservO.WwithRD", "iL", "iXp", "iYear")),
             adjusted = build_formula("iGDP", list("iKservO.WwithRD", "ihLest", "iU", "iYear")))
  )
@

<<Data>>=
# Create the historicalData data frames
historicalData <- list()
historicalData$PT <- filter(IST, Country == "PT")
historicalData$UK <- filter(Leeds, Country == "UK")
allHistoricalData <- do.call(plyr::rbind.fill, historicalData)
@

@@

\begin{table} \caption{Unadjusted time-series variables.} 
\label{tab:unadjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal           & UK                 \\
      \midrule
      Output      & \Sexpr{summands(formulas$PT$unadjusted, n=1, right=FALSE)}   
                                                             & \Sexpr{summands(formulas$UK$unadjusted, n=1, right=FALSE)}\\
      Capital     & \Sexpr{summands(formulas$PT$unadjusted, 1)} & \Sexpr{summands(formulas$UK$unadjusted, 1)} \\
      Labor       & \Sexpr{summands(formulas$PT$unadjusted, 2)} & \Sexpr{summands(formulas$UK$unadjusted, 2)} \\
      Energy      & \Sexpr{summands(formulas$PT$unadjusted, 3)} & \Sexpr{summands(formulas$UK$unadjusted, 3)} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{Quality-adjusted time series variables.} 
\label{tab:quality_adjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal         & UK               \\
      \midrule
      Output      & \Sexpr{summands(formulas$PT$adjusted, n=1, right=FALSE)}   
                                                           & \Sexpr{summands(formulas$UK$adjusted, n=1, right=FALSE)} \\
      Capital     & \Sexpr{summands(formulas$PT$adjusted, 1)} & \Sexpr{summands(formulas$UK$adjusted, 1)} \\
      Labor       & \Sexpr{summands(formulas$PT$adjusted, 2)} & \Sexpr{summands(formulas$UK$adjusted, 2)} \\
      Energy      & \Sexpr{summands(formulas$PT$adjusted, 3)} & \Sexpr{summands(formulas$UK$adjusted, 3)} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<Historical Data, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Historical data.", warning=FALSE>>=
histDataForGraph <- leaf_apply(formulas, class = "formula", f=function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]
  measure.vars <- c(summands(form, right=FALSE), summands(form, 1:3))
  trimmed_historical_data <- historicalData[[Country]][c("Year", measure.vars)]
  mtrimmed_historical_data <- melt(trimmed_historical_data, measure.vars = measure.vars, 
                                   value.name = "value", variable.name = "variable")
  mtrimmed_historical_data$Country <- Country
  if (flavor == "unadjusted"){
    mtrimmed_historical_data$flavor <- "Unadjusted"
  } else if (flavor == "adjusted") {
    mtrimmed_historical_data$flavor <- "Quality-adjusted"
  } else {
    stop(paste("Unknown flavor", flavor, "in histDataForGraph."))
  }
  return(mtrimmed_historical_data)
}) %>% do.call(rbind, .)

# Relevel factors before plotting
histDataForGraph$Country <- relevelFactor(as.factor(histDataForGraph$Country), c("PT", "UK"))
histDataForGraph$flavor <- relevelFactor(as.factor(histDataForGraph$flavor), c("Unadjusted", "Quality-adjusted"))
histDataForGraph$variable <- relevelFactor(as.factor(histDataForGraph$variable), unique(histDataForGraph$variable))

historicalPlot(histDataForGraph, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = Country ~ flavor, 
               line_types = c(1,2,4,3,2,4,3,2,3,3)) + 
  scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
  coord_cartesian(ylim=c(0.5, 13)) +
  xlab("") + 
  ylab("Indexed (1960=1)")
@

<<Pre-econometric graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Pre-econometric data.", warning=FALSE>>=

# Create some pre-econometric ratios

# # Melt the data
# mdata_preeconometric <- melt(historicalData, 
#                              measure.vars = c("kPerGDP.unadj", "lPerGDP.unadj", "ePerGDP.unadj",
#                                               "kPerGDP.adj", "lPerGDP.adj", "ePerGDP.adj"),
#                              variable.name = "variable",
#                              value.name = "value")
# # Add a column to say whether the variable is Unadjusted or Quality-adjusted based on variable name
# mdata_preeconometric$adjusted <- lapply(mdata_preeconometric$variable, function(x) {
#   suffix <- unlist(strsplit(as.character(x), split="[.]")) [2]
#   if (suffix == "unadj"){
#     return("Unadjusted")
#   } else if (suffix == "adj"){
#     return("Quality-adjusted")
#   } else {
#     stop(paste("Unknown suffix:", suffix))
#   }
# })

pre_econometric_data <- leaf_apply(formulas, class = "formula", f=function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]
  nums <- summands(form, n=1:3)
  denom <- summands(form, n=1, right=FALSE)
  ratios <- lapply(nums, function(num) {paste(num, denom, sep="/")}) %>% unlist
  coltemplate <- build_formula(NULL, c("Year", ratios))
  trimmed_historical_data <- historicalData[[Country]][all.vars(coltemplate)]
  ratio_data <- evalFormula(coltemplate, data=trimmed_historical_data)$right
  mratio_data <- melt(ratio_data, measure.vars = ratios, value.name = "ratio", variable.name = "variable")
  mratio_data$Country <- Country
  mratio_data$flavor <- flavor
  return(mratio_data)
}) %>% do.call(rbind, .)

# Fix levels
pre_econometric_data$flavor <- relevelFactor(as.factor(as.character(pre_econometric_data$flavor)), 
                                               c("Unadjusted", "Quality-adjusted"))
pre_econometric_data$Country <- relevelFactor(as.factor(pre_econometric_data$Country), 
                                              c("PT", "UK"))
pre_econometric_data$variable <- relevelFactor(as.factor(pre_econometric_data$variable),
                                               unique(pre_econometric_data$variable))
# Now make the plot
historicalPlot(pre_econometric_data, 
               mapping=aes(x=Year, y=ratio, group=variable, linetype=variable), 
               facet_formula = Country ~ flavor, 
               line_types = c(2,4,3, 2,4,3, 2,3,3)) + 
  xlab("") + 
  ylab("Indexed (1960=1)")
@

@@

#   models <- list()
#   for ()
